Individual and Differential Harm in Redistricting

### 2022-12-06 paper/harm.Rmd
fix fig ref
anon

---
title: Individual and Differential Harm in Redistricting
authors:
  - name: Cory McCartan
    thanks: |
      To whom correspondence should be addressed.
      The authors thank Steve Ansolabehere, Matt Blackwell, Kosuke Imai, Gary King, 
      Shiro Kuriwaki, Tyler Simko, Nick Stephanopolous, and 
      participants of the Fall 2021 Harvard American Politics Research Workshop,
      Fall 2022 Harvard Applied Statistics Workshop, 
      and Imai Research Group for helpful comments and discussion.
      Replication code and data are available at <https://github.com/CoryMcCartan/harm-redistricting>.
    department: Ph.D. Candidate, Department of Statistics
    affiliation: Harvard University
    location: 1 Oxford St, Cambridge, MA
    email: cmccartan@fas.harvard.edu
  - name: Christopher T. Kenny
    department: Ph.D. Candidate, Department of Government
    affiliation: Harvard University
    location: 1737 Cambridge St., Cambridge, MA
    email: christopherkenny@fas.harvard.edu
abstract: |
  Social scientists have developed dozens of measures for assessing partisan bias in redistricting.
  But these measures are not easily adapted to other groups, 
  including groups defined by race, class, or geography.
  Nor are they applicable to single- or no-party contexts, such as local redistricting.
  To overcome these limitations, we propose a unified framework of harm for evaluating 
  the impacts of a districting plan on individual voters and the groups to which they belong. 
  We consider a voter harmed if their chosen candidate is not elected under the current plan, 
  but would be under a different plan.
  Harm improves on existing measures by both focusing on the choices of individual voters 
  and directly incorporating counterfactual plans.
  We discuss strategies for estimating harm, and demonstrate the utility of our framework 
  through analyses of partisan gerrymandering in New Jersey, voting rights 
  litigation in Alabama, and racial dynamics of Boston City Council elections.
keywords:
  - representation
  - redistricting
  - voting rights
  - individual harm
bibliography: references.bib
biblio-style: apsr
output:
  rticles::arxiv_article:
    keep_tex: true
    includes:
      in_header: "header.tex"
    latex_engine: pdflatex
fontsize: 12pt
editor_options:
  markdown:
    wrap: sentence
---

@@

Figure \@ref(fig:alabama-harm-sample) summarizes the results, which are mostly similar to the results in in Section \ref{sec:alabama}.
Across the state, the estimated average harm is 6.9%, which translates to around 144,000 voters.
Of these harmed voters, around 56,000 are Black, while 88,000 are White.
Similarly, 73,000 are Democratic, while only 71,000 are Republican.
Normalizing by group sizes, the expected harm for a Black Democrat is 9.6%, for a Black Republican is 16.8%, for a White Democrat is 9.7%, and for a White Republican is 5.2%.
So the challenged plan differentially harms Democratic voters, but does not differentially harm Black Democrats versus White Democrats.
However, it it especially harms Black Republicans, who have an estimated differential harm of 7.2% versus Black Democrats and of 11.6% versus White Republicans.

>>>>

---
title: Individual and Differential Harm in Redistricting
authors:
  - name: Anonymized Author(s)
    affiliation: Anonymized Institution
abstract: |
  Social scientists have developed dozens of measures for assessing partisan bias in redistricting.
  But these measures are not easily adapted to other groups, 
  including groups defined by race, class, or geography.
  Nor are they applicable to single- or no-party contexts, such as local redistricting.
  To overcome these limitations, we propose a unified framework of harm for evaluating 
  the impacts of a districting plan on individual voters and the groups to which they belong. 
  We consider a voter harmed if their chosen candidate is not elected under the current plan, 
  but would be under a different plan.
  Harm improves on existing measures by both focusing on the choices of individual voters 
  and directly incorporating counterfactual plans.
  We discuss strategies for estimating harm, and demonstrate the utility of our framework 
  through analyses of partisan gerrymandering in New Jersey, voting rights 
  litigation in Alabama, and racial dynamics of Boston City Council elections.
keywords:
  - representation
  - redistricting
  - voting rights
  - individual harm
bibliography: references.bib
biblio-style: apsr
output:
  rticles::arxiv_article:
    keep_tex: true
    includes:
      in_header: "header.tex"
    latex_engine: pdflatex
fontsize: 12pt
editor_options:
  markdown:
    wrap: sentence
---

@@

Figure \ref{fig:alabama-harm-sample} summarizes the results, which are mostly similar to the results in in Section \ref{sec:alabama}.
Across the state, the estimated average harm is 6.9%, which translates to around 144,000 voters.
Of these harmed voters, around 56,000 are Black, while 88,000 are White.
Similarly, 73,000 are Democratic, while only 71,000 are Republican.
Normalizing by group sizes, the expected harm for a Black Democrat is 9.6%, for a Black Republican is 16.8%, for a White Democrat is 9.7%, and for a White Republican is 5.2%.
So the challenged plan differentially harms Democratic voters, but does not differentially harm Black Democrats versus White Democrats.
However, it it especially harms Black Republicans, who have an estimated differential harm of 7.2% versus Black Democrats and of 11.6% versus White Republicans.
